\newpage
```{r, include = FALSE}
knitr::opts_chunk$set(
  dpi = 600, dev = "jpeg", collapse=TRUE, options(digits=4)
)
# options(knitr.duplicate.label = "allow")
```

# Introduction: Experiments and Variables 

## Chapter pre-cap

## Experiments and effects {#c1-exp-and-effects}

### Experiments and inference {#c1-exp-and-inference}

## Our experiment {#c1-exp}

### Our experiment: Introduction {#c1-exp-intro}
 
```{r F1-1, echo = FALSE, fig.height = 2.75, fig.width = 8, echo = FALSE, fig.cap = "(left) Average height of males and females in the United States of America, organized by age (Fryar et al. (2012)). (middle) Average f0 produced by male and female speakers from 5 years of age until adulthood (Lee et al. 1999). (right) Average acoustic vocal-tract length (VTL) of male and female speakers from 5 years of age until adulthood (Lee et al. 1999). The average adult male VTL was set to 15 cm. Error bars indicate two standard deviations.", eval = TRUE}

################################################################################
### Figure 1.1
################################################################################
library (bmmb)
data (height_data)
data (lee_etal_data)

par (mfrow = c(1,3), mar = c(4.1,4.1,1,1))
plot (height_data$age[height_data$gender=="f"]-.1,
      height_data$height[height_data$gender=="f"],
      pch=16,col=2,lwd=2,cex=1.5, ylim = c(70,205),type='b', xlab="Age (years)",
      ylab = "Height (cm)",xlim=c(2,21),cex.axis=1.3,cex.lab=1.3)
lines (height_data$age[height_data$gender=="m"]+.1,
       height_data$height[height_data$gender=="m"],
      pch=16,col=4,lwd=2,cex=1.5,type='b')
grid()
legend (13,110,legend = c("Female","Male"), col = c(2,4),pch=16,cex=1.2,
        pt.cex=1.5, bty='n')

phonTools::errorbar(height_data$age[height_data$gender=="f"]-.1,
                    height_data$height[height_data$gender=="f"],
                    height_data$sd[height_data$gender=="f"]*2,col=2,lwd=1,length=0.051)
phonTools::errorbar(height_data$age[height_data$gender=="m"]+.1,
                    height_data$height[height_data$gender=="m"],
                    height_data$sd[height_data$gender=="m"]*2,col=4,lwd=1,length=0.051)

rect (9.5,120,12.5,177,lwd=2,border="forestgreen",lty=1)
rect (17.5,143,20.5,205,lwd=2,border="forestgreen",lty=1)


lee_etal_data$f0_original=lee_etal_data$f0

lee_etal_data$f0 = log(lee_etal_data$f0_original)
aggdmu = aggregate (f0 ~ age+gender, FUN = mean, data = lee_etal_data)
aggdmu[,3] = exp(aggdmu[,3])

lee_etal_data$f0 = log(lee_etal_data$f0sd/lee_etal_data$f0_original)
aggdsd = aggregate (f0 ~ age+gender, FUN = sd, data = lee_etal_data)
aggdsd[,3] = (exp(aggdsd[,3])-1)*aggdmu[,3] 

aggdmuf0 = aggdmu
aggdsdf0 = aggdsd

plot (aggdmu$age[aggdmu$gender=="f"]-.1,aggdmu$f0[aggdmu$gender=="f"],
      pch=16,col=2,lwd=2,cex=1.5, ylim = c(75,370),type='b', xlab="Age (years)",
      ylab = "f0 (Hz)",xlim=c(2,21),cex.axis=1.3,cex.lab=1.3)
lines (aggdmu$age[aggdmu$gender=="m"]+.1,aggdmu$f0[aggdmu$gender=="m"],
      pch=16,col=4,lwd=2,cex=1.5,type='b')
grid()

phonTools::errorbars(aggdmu$age[aggdmu$gender=="f"]-.1,aggdmu$f0[aggdmu$gender=="f"],
                   aggdsd$f0[aggdsd$gender=="f"]*2,col=2,lwd=1,length=0.051)
phonTools::errorbars(aggdmu$age[aggdmu$gender=="m"]+.1,aggdmu$f0[aggdmu$gender=="m"],
                   aggdsd$f0[aggdsd$gender=="m"]*2,col=4,lwd=1,length=0.051)


gbars = (log(lee_etal_data$f1)+log(lee_etal_data$f2)+log(lee_etal_data$f3))/3

aggdmu = aggregate (gbars ~ age+gender, FUN = mean, data = lee_etal_data)
aggdmu[,3] = (aggdmu[,3])

aggdmu$vtl = exp(-((aggdmu$gbar)-min(aggdmu$gbars)))
aggdmu$vtl = 15 * aggdmu$vtl

vtl = (log(lee_etal_data$f1sd/lee_etal_data$f1)+
         log(lee_etal_data$f2sd/lee_etal_data$f2)+
         log(lee_etal_data$f3sd/lee_etal_data$f3))/3

aggdsd = aggregate (vtl ~ age+gender, FUN = mean, data = lee_etal_data)
aggdsd[,3] = exp(aggdsd[,3])*aggdmu$vtl

aggdmuvtl = aggdmu
aggdsdvtl = aggdsd

plot (aggdmu$age[aggdmu$gender=="f"]-.1,aggdmu$vtl[aggdmu$gender=="f"],
      pch=16,col=2,lwd=2,cex=1.5, ylim = c(8.5,17),type='b', xlab="Age (years)",
      ylab = "VTL (cm)",xlim=c(2,21),cex.axis=1.3,cex.lab=1.3)
lines (aggdmu$age[aggdmu$gender=="m"]+.1,aggdmu$vtl[aggdmu$gender=="m"],
      pch=16,col=4,lwd=2,cex=1.5,type='b')
grid()

phonTools::errorbars(aggdmu$age[aggdmu$gender=="f"]-.1,
                   aggdmu$vtl[aggdmu$gender=="f"],
                   aggdsd$vtl[aggdsd$gender=="f"]*2,col=2,lwd=1,length=0.051)
phonTools::errorbars(aggdmu$age[aggdmu$gender=="m"]+.1,
                   aggdmu$vtl[aggdmu$gender=="m"],
                   aggdsd$vtl[aggdsd$gender=="m"]*2,col=4,lwd=1,length=0.051)

```

### Our experimental methods {#c1-methods}

### Our research questions {#c1-research-questions}

### Our experimental data {#c1-exp-data}

```{r, collapse = TRUE, messages=FALSE,warning=FALSE}
# load package
library ("bmmb")

# load data
data (exp_data_all)
```

```{r, collapse = TRUE}
# see first 6 rows
head (exp_data_all)
```

```{r, collapse = TRUE, eval = FALSE}
write.csv (exp_data_all, "exp_data_all.csv", row.names = FALSE)
```

```{r, collapse = TRUE}
str (exp_data_all)
```

```{r, collapse = TRUE, eval = FALSE}
exp_data_all$height
```

```{r, collapse = TRUE}
# show the first six
head(exp_data_all$height)

# show the first element
exp_data_all$height[1]

# show elements 2 to 6
exp_data_all$height[2:6]
```

```{r, collapse = TRUE}
head(exp_data_all[[3]])

head(exp_data_all[["height"]])
```

```{r, collapse = TRUE}
head(exp_data_all[,3])

head(exp_data_all[,"height"])
```

```{r, collapse = TRUE}
exp_data_all[1,]

exp_data_all[1,2]
```

## Variables {#c1-variables}

### Populations and samples {#c1-pops-and-samps}

### Dependent and Independent Variables {#c1-dep-and-indep}

### Categorical variables and 'factors' {#c1-categorical}

```{r, collapse = TRUE}
# see the first 6 observations
head (exp_data_all$C_v)   

# it has levels
levels(exp_data_all$C_v)  

# each level has numerical values
table (exp_data_all$C_v, as.numeric (exp_data_all$C_v))  
```

```{r, collapse = TRUE}
# re order
exp_data_all$C_v_f = factor (exp_data_all$C_v, levels = c('w','m','g','b'))

# the new order is evident
levels (exp_data_all$C_v_f)

# note that 'm' is now the second category
xtabs ( ~ exp_data_all$C_v + exp_data_all$C_v_f)
```

### Quantitative variables  {#c1-quantitative}

```{r, collapse = TRUE}
2 == 1

"hello" == "hello"

"hello" != "hello"
```

```{r, collapse = TRUE, eval = TRUE}
2 > 1

2 >= 1

2 < 1

2 >= 1
```

```{r, collapse = TRUE}
TRUE + 1

(2 == 2) + 1
```

```{r, collapse = TRUE}
# are the values less than or equal to 3?
c(1,2,3,4,5,6,7,8,9,10) <= 3
```

```{r, collapse = TRUE}
logical_vector = c(1,2,3,4,5,6,7,8,9,10) <= 3

as.numeric (logical_vector)

sum (logical_vector)

sum (c(1,2,3,4,5,6,7,8,9,10) <= 3)
```

```{r, collapse = TRUE}
# TRUE if f0 < 175
f0_idx = exp_data_all$f0 < 175

str (f0_idx)

sum (f0_idx)
```

```{r, collapse = TRUE}
# get only rows where f0 < 175, i.e. where f0_idx is TRUE
low_f0 = exp_data_all[f0_idx,]

nrow(low_f0)

max(low_f0$f0)
```

```{r, collapse = TRUE}
# get only rows where f0 >= 175, i.e. where f0_idx is FALSE
high_f0 = exp_data_all[!f0_idx,]

nrow(high_f0)

min(high_f0$f0)
```

## Inspecting our data {#c1-inspecting}

### Inspecting categorical variables {#c1-inspecting-categorical}

```{r, collapse = TRUE}
table (exp_data_all$L)
```

```{r, collapse = TRUE}
# table of listener and veridical speaker category
table (exp_data_all$C_v, exp_data_all$L)
```

```{r, collapse = TRUE}
# table of listener and apparent speaker category
table (exp_data_all$C, exp_data_all$L)
```

```{r F1-2, fig.height = 2.5, fig.width = 8, echo = FALSE, fig.cap = "Comparisons of mosaic plots showing variables that do not (left), and do (right), affect each other.",collapse=TRUE}

################################################################################
### Figure 1.2
################################################################################

par (mfrow = c(1,2), mar=c(2,2,1,.5))
plot (t(table (exp_data_all$C_v, exp_data_all$L)), main = '',xlab='Listener', 
      col = bmmb::cols[2:5],ylab = 'Veridical Speaker Class')
plot(t(table (exp_data_all$C, exp_data_all$L)), main = '',xlab='Listener', 
     col = bmmb::cols[2:5],ylab = 'Apparent Speaker Class')
```

```{r, eval = FALSE,collapse=TRUE}
tmp_tab = table (exp_data_all$C, exp_data_all$L, exp_data_all$R)
tmp_tab
tmp_tab[,,1]
tmp_tab[,,2]
```

```{r F1-3, fig.height = 2.5, fig.width = 8, echo = FALSE, fig.cap = "Mosaic plots highlighting a three-way relationship: The two-way listener by apparent speaker class relationship varies as a function of the third variable, resonance (indicated along the top of each plot).", echo=FALSE}

################################################################################
### Figure 1.3
################################################################################

tmp_tab = table (exp_data_all$C, exp_data_all$L, exp_data_all$R)

par (mfrow = c(1,2), mar=c(2,2,2.5,0.5))
plot (t(tmp_tab[,,1]), main = 'Actual Resonance',xlab='Listener', 
      col = bmmb::cols[2:5],
      ylab = 'Apparent Speaker Class')
plot(t(tmp_tab[,,2]), main = 'Big Resonance',xlab='Listener', 
     col = bmmb::cols[2:5],
     ylab = 'Apparent Speaker Class')
```

### Inspecting quantitative variables {#c1-inspecting-quantitative}   

```{r, collapse = TRUE}
# calculate the mean
mean (exp_data_all$height)

# find the number of observations
length (exp_data_all$height)

# find quantiles
quantile (exp_data_all$height)
```

```{r F1-4, fig.height = 4.5, fig.width = 8, fig.cap='Each column presents the same data in three different ways: (top) As individual jittered points, (middle) as a boxplot, and (bottom) as a histogram.', echo = FALSE}

################################################################################
### Figure 1.4
################################################################################

mens_height = exp_data_all$height[exp_data_all$C_v=='m']

set.seed(7)
par (mfcol = c(3,2), mar = c(0.5,4,0.5,1), oma = c(4,0,2,0))
plot (mens_height, jitter (rep(1,length(mens_height))), xlim = c(100, 210), ylim = c(.95,1.05),
      yaxt='n',ylab='', pch = 16, col = yellow, xaxt='n')
mtext (side =3, outer = FALSE, text = "Adult male apparent heights", line = 1)
boxplot (mens_height, horizontal = TRUE, ylim = c(100, 210), col = coral,xaxt='n')
hist (mens_height,main="", col = teal, xlim = c(100, 210),breaks=40,cex.lab=1.3,
      cex.axis=1.3, xlab = "")
mtext (side =1, outer = FALSE, text = "Apparent height (cm)", line = 3)
box()

plot (exp_data_all$height, jitter (rep(1,length(exp_data_all$height))), xlim = c(100, 210), ylim = c(.95,1.05),
      yaxt='n',ylab='', pch = 16, col = yellow, xaxt='n')
mtext (side =3, outer = FALSE, text = "All apparent heights", line = 1)
boxplot (exp_data_all$height, horizontal = TRUE, ylim = c(100, 210), col = coral,xaxt='n')
hist (exp_data_all$height,main="", col = teal, xlim = c(100, 210),breaks=40,cex.lab=1.3,
      cex.axis=1.3,xlab = "")
mtext (side =1, outer = FALSE, text = "Apparent height (cm)", line = 3)

box()
```


```{r F1-5, fig.height = 5, fig.width = 8, fig.cap='A pairs plot of the continuous variables in our data, showing different sorts of relationships between: f0 (fundamental frequency in Hertz), vtl (vocal-tract length in centimeters), dur (duration in milliseconds), and height (apparent speaker height in centimeters).', echo = FALSE}

################################################################################
### Figure 1.5
################################################################################

agg_data = aggregate (height~f0+vtl+dur, data = exp_data_all, FUN=mean)
pairs (agg_data, col = lavender,pch=16)

```


### Exploring continuous and categorical variables together {#c1-inspecting-together}

```{r, eval = FALSE}
boxplot (y ~ factor)
```

```{r F1-6, fig.height = 2.5, fig.width = 8, fig.cap='Boxplots showing the distribution of different quantitative variables in our data according to the veridical speaker categories of boy (b), girl (g), man (m), and woman (w).', echo = FALSE}

################################################################################
### Figure 1.6
################################################################################

par (mfrow = c(1,3), mar = c(3,4.2,1,1))
boxplot (f0 ~ C_v, data = exp_data_all, col = bmmb::cols[1:4],cex.lab=1.3,cex.axis=1.3,
         xlab="",ylab="f0 (Hz)")
boxplot (vtl ~ C_v, data = exp_data_all, col = bmmb::cols[5:8],cex.lab=1.3,cex.axis=1.3,
         xlab="",ylab="Vocal-tract Length (cm)")
boxplot (dur ~ C_v, data = exp_data_all, col = bmmb::cols[9:12],cex.lab=1.3,cex.axis=1.3,
         xlab="",ylab="Duration (ms)")
```

```{r F1-7, fig.height = 2.5, fig.width = 8, fig.cap='Speakers plotted according to their fundamental frequency (f0), vocal-tract length, and average apparent height. Letters indicate if speaker is a boy (b), girl (g), man (m), or woman (w).', echo = FALSE}

agg_data = aggregate (cbind(f0,vtl,dur,height)~C_v+S, data = exp_data_all, FUN=mean)


par (mfrow = c(1,3), mar = c(4,4,1,1))

plot (agg_data$vtl, agg_data$height, type = 'n',ylab="Apparent height (cm)",xlab="Vocal-tract length (cm)",
      cex.lab=1.3,cex.axis=1.3)
text (agg_data$vtl, agg_data$height, labels = agg_data$C_v, 
      col = bmmb::cols[2:5][factor(agg_data$C_v)], cex = 1.5)

plot (agg_data$f0, agg_data$height, type = 'n',xlab="f0 (Hz)",ylab="Apparent height (cm)",
      cex.lab=1.3,cex.axis=1.3)
text (agg_data$f0, agg_data$height, labels = agg_data$C_v, 
      col = bmmb::cols[2:5][factor(agg_data$C_v)], cex = 1.5)

plot (agg_data$vtl, agg_data$f0, type = 'n',ylab="f0 (Hz)",xlab="Vocal-tract length (cm)",
      cex.lab=1.3,cex.axis=1.3)
text (agg_data$vtl, agg_data$f0, labels = agg_data$C_v, 
      col = bmmb::cols[2:5][factor(agg_data$C_v)], cex = 1.5)
```


## Exercises

## References

## Plot Code

```{r , echo = FALSE}
labs = knitr::all_labels()
labs = labs[grep ("F", labs)]
```

```{r , ref.label=labs, eval=FALSE}
```
